---
title: "Week 8b: Functional Programming Lab"
author: "Emily Wilson"
date: "3/24/2021"
output:
  rmdformats::material
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# Load libraries

* **tidyverse**
* **PNWColors** for color palettes
* **palmerpenguins**  for data
* **kableExtra** for pretty tables

```{r}
library(tidyverse)
library(PNWColors)
library(palmerpenguins)
library(kableExtra)
```

# Write a Function

### Write a function to create a ratio of different penguin body measurements

```{r}
measurement_ratio <- function(x, y) { # name the function and give it an input
  ratio <- (x / y) # put in the equation I want it to use
  return(ratio) # tell it what values I want returned
}
```

### Test the function

We tested this function by pairing it with mutate() to create additional columns that gave us the ratio of bill length (mm) to bill depth (mm), a ratio of body mass (g) to bill depth (mm), and a ratio of body mass (g) to bill length (mm)

```{r}
penguins %>% # choose data
  drop_na() %>% # remove NA values
  mutate(length_depth_ratio = measurement_ratio(bill_length_mm, bill_depth_mm), # test the function with a ratio of bill length to bill depth
         mass_depth_ratio = measurement_ratio(body_mass_g, bill_depth_mm), # test the function with a ratio of body mass to bill depth
         mass_length_ratio = measurement_ratio(body_mass_g, bill_length_mm)) %>% # test the function with a ratio of body mass to bill length
  select(species, island, length_depth_ratio, mass_depth_ratio, mass_length_ratio) %>% # show only species, island, and our converted columns
  kbl(col.names = c("Species", "Island", "Length:Depth", "Mass:Depth", "Mass:Length"), align = "c") %>% # tidy column titles
  kable_styling(bootstrap_options = c("responsive", "condensed"), font_size = 12) %>% 
  scroll_box(width = "500px", height = "200px")
```

# Write a Graph Function

### Write a function that creates a graph...

```{r}
penguinplot <- function(data = penguins, x, y, z, cat = TRUE){ # name the function, tell it to use penguin data and give it inputs
  pal <- pnw_palette("Starfish", 3, type = "discrete")# give a color palette to use
  
  if(cat == TRUE) {ggplot(data, aes(x = {{x}}, y = {{y}} , color = {{z}})) + # if cat == TRUE, I want a boxplot
    geom_boxplot() +
    scale_color_manual("Species", values=pal) + 
    theme_classic()
  } 
  
else{ # if cat == FALSE, I want a scatterplot
  ggplot(data, aes(x = {{x}}, y = {{y}} , color = {{z}})) +
    geom_point() +
    geom_smooth(method = "lm") +
    scale_color_manual("Species", values=pal) + 
    theme_classic()
}
}
```


### Test the function

### Get a boxplot
```{r}
penguinplot(x = island, y = body_mass_g, z = species) # get a boxplot (cat == TRUE by default)
```


### Get a scatterplot
```{r}
penguinplot(x = body_mass_g, y = bill_length_mm, z = species, cat = FALSE) # get a scatterplot (cat == FALSE)
```











