---
title: "Week 7a Map Lab"
author: "Emily Wilson"
date: "3/8/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    toc: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, # 
                      warning = FALSE, # don't show any warnings in final product
                      message = FALSE, # don't show any messages in final product
                      fig.path ="../outputs/") # put all figures into outputs folder
```


# Introduction

This is a document for Computer Modeling Week 7a lab to explore making maps in R. The data set for this document can be found at the [Plastic Pollution](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-26/readme.md) page from R for Data Science's Tidy Tuesday repository.

***

# Load libraries

* *tidyverse*
* *here*
* *tidytuesdayr*
* *maps*
* *mapdata*
* *mapproj*

```{r libraries, include = FALSE}
library(tidytuesdayR)
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(giphyr)
```

***

# Read in & process data

* This plastics data comes from **"Break Free of Plastic"**, courtesy of Sarah Sauve.
  * This data was used for Tidy Tuesday for the week of 2021-01-26.  
  
* World map data was obtained using the *{maps}* package in R.

* Plastics data was joined with world data to visualize the amount of plastic collected in each country.

```{r data, include = FALSE}
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv') # Tidy Tuesday plastics data
world <- map_data("world") # data for the world
```


```{r manipulation, include = FALSE}
# summarize data so all we have is total 2019 plastic pollution by country
plastic_summ <- plastics %>% 
  rename(region = country) %>% # change column name so it matches with the world data set
  filter(year == "2019", parent_company == "Grand Total")# retain only data on grand totals of plastics per country for 2019 
  
plastic_summ$region[13] = "Ivory Coast" # rename all of the weirdness in the plastics data set
plastic_summ$region[15] = "Ecuador"
plastic_summ$region[33] = "Nigeria"
plastic_summ$region[43] = "Taiwan"
plastic_summ$region[50] = "UK"
plastic_summ$region[51] = "USA"

 plastic_summ <- plastic_summ %>% 
  right_join(world) %>% # join with world data
  select(region, grand_total, long, lat, group) # only keep columns needed for the map
```

***

# Plastic pollution map

```{r plasticsmap, echo = FALSE, fig.align = 'center', fig.cap=  "Plastic pollution by country in 2019. Countries in gray had no data available."}
ggplot() +
  geom_polygon(data = plastic_summ,
               aes(x = long, # longtitude
                   y = lat, # latitude
                   group = group,  # keep it from being an abstract art piece
                   fill = grand_total)) + # fill by plastic pollution
  coord_map(projection = "mercator", xlim = c(-180,180)) + # set projection and x lims
  scale_fill_viridis_c(trans = "log10") + # make the scale fill logarithmic for easier interpretation
  theme_void() + # change theme
  labs(fill = "log(plastics)", # change to scale name to represent what it's actually showing
       caption = "Source: Break Free of Plastic") + 
  theme(panel.background = element_rect(fill = "lightcyan"), # make the ocean areas blue
        panel.grid.major = element_blank()) # remove annoying grid lines

```
